Manuscript draft

### 2019-05-30 manuscript/manuscript draft.Rmd
updated manuscript draft

*NPP* The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests, with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].

@@

*[insert table here similar to in trop-forC paper, including information on productivity measures, methods used to estimate, components included etc]*

Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. Sites that had experienced significant disturbance were also excluded. Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.  
There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems [@DeLucia2007; @Gillman2015]. To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.

>>>>

*NPP* The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests [@Huston2009], with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].

Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP. Studies on global patterns of ANPP are equally inconclusive: Gillman et al. (2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands. These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].

@@

| Variable | Definition | Components included | Methodologies used|
|-------------|-------------------|-------------|------------------------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and optionally understory | Direct measurement of annual increments of components | 
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 
| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |

Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. Sites that had experienced significant disturbance were also excluded. Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.  
There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015]. To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
